\name: A hierarchical Bayesian model for Supernova Cosmology

### 2018-02-23 papers/methods/paper.tex
Abstract

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\name}{SN-BHM}
\newcommand{\myemail}{samuelreay@gmail.com}
\newcommand\abs[1]{\left|#1\right|}
\newcommand {\etal} {\emph{~et~al.} }
\newcommand{\green}{\color{green}}
\newcommand{\blue}{\color{blue}}
\newcommand{\red}{\color{red}}

@@

% Abstract of the paper
\begin{abstract}
Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract Abstract
\end{abstract}

@@

In this paper, we lay out a new hierarchical model that advances the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects whilst improving computational performance. Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology. In Section \ref{sec:method} we outline our methodology. Model verification on simulated datasets is given in Section \ref{sec:verification}. Forecasts for the impending DES three year spectroscopic supernova survey are contained in Section \ref{sec:forecasts}. {\red Section \ref{sec:sys} investigates the effect of various systematics on our model, and Section \ref{sec:details} provides details on potential areas of improvement and unsuccessful methodologies. }

@@

The primary numerical improvement we made on existing frameworks was to remove costly probability evaluations of multivariate normals. To increase efficiency, the optimum way to sample a multivariate normal is to reparameterise it such that instead of sampling $\mathcal{N}(\vec{x}|\vec{\mu}, \Sigma)$, you sample $\mathcal{N}(\vec{\delta}|0,1)$ where $\vec{x} = \vec{\mu} + L \vec{\delta}$ and $L$ is the cholesky decomposition of $\Sigma$. In this way, we can efficiently sample the unit normal probability distribution instead of sampling a multivariate normal probability distribution. Switching to this parametrisation resulted in a computational increase of an order of magnitude, taking fits for a sample of approximately 500 supernovae from roughly four hours down to twenty minutes.

>>>>

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\name}{Steve-o}
\newcommand{\myemail}{samuelreay@gmail.com}
\newcommand\abs[1]{\left|#1\right|}
\newcommand {\etal} {\emph{~et~al.} }
\newcommand{\green}{\color{green}}
\newcommand{\blue}{\color{blue}}
\newcommand{\red}{\color{red}}

@@

% Abstract of the paper
\begin{abstract}
We present a hierarchical Bayesian model for use in supernova cosmology which takes into account full analysis systematics, Malmquist bias, underlying populations and standardisation coefficients. Our model is optimised for efficient fitting and has been validated on a combined dataset of more than $250\,000$ supernovae. Simplified statistical simulations for ideal supernovae cosmology show no bias, and sophisticated SNANA simulations quantify the bias for Flat $w$CDM with a matter prior $\Omega_m \sim \mathcal{N}(0.3, 0.01)$ allow rigorous determination of model bias in $w$ when provided with information on matter density (such as might be found from CMB constraints). Bias is found to range from $-0.15\sigma$ to $+0.27\sigma$ across intrinsic dispersion models for datasets of comparable strength to the Dark Energy Survey three-year analysis. 
\end{abstract}

@@

In this paper, we lay out a new hierarchical model that advances the past work of \citet{Rubin2015} to include more robust treatment of systematics and selection effects whilst improving computational performance. Section \ref{sec:review} is dedicated to a quick review of the supernovae cosmology. In Section \ref{sec:method} we outline our methodology. Model verification on simulated datasets is given in Section \ref{sec:verification}, along with details on potential areas of improvement.

@@

The primary numerical improvement we made on existing frameworks was to remove costly probability evaluations of multivariate normals. To increase efficiency, the optimum way to sample a multivariate normal is to reparameterise it such that instead of sampling $\mathcal{N}(\vec{x}|\vec{\mu}, \Sigma)$, you sample $\mathcal{N}(\vec{\delta}|0,1)$ where $\vec{x} = \vec{\mu} + L \vec{\delta}$ and $L$ is the cholesky decomposition of $\Sigma$. In this way, we can efficiently sample the unit normal probability distribution instead of sampling a multivariate normal probability distribution. Switching to this parametrisation resulted in a computational increase of an order of magnitude, taking fits for a sample of approximately 500 supernovae from roughly four hours down to thirty minutes.
